### 2009-07-30 tags/ant-crisp/Papers/ci-spark/experiments.tex
Tagged the last version of CRISP with Ant buildfile and directory structure before converting to Maven.

We now return to the original question of the paper: is planning technology
ready for realistic applications in natural language generation? To investigate
this question we consider two sets of experiments, designed to evaluate the
performance of several planners on the NLG planning domains from the previous
section. Starting with the CRISP domain, we first present a scenario which
focuses on the generation of referring expressions with a tiny grammar
(Section~\ref{sec:exper-1:-sent}). We then look at a setting in which CRISP is
used for surface realization with the XTAG Grammar \citep{xtag01:_tr}, a
large-scale TAG grammar for English (Section~\ref{sec:exper-2:-sent-xtag}). In
the second set of experiments we investigate the GIVE domain. We begin with a
domain that is similar to the classic Gridworld
(Section~\ref{sec:exper-3:-minim}), and then add extra grid cells to the world
that are not necessary to complete the task
(Section~\ref{sec:experiment-4:-give}). We also investigate the role that goal
ordering plays in these problems. These experiments are configured in a way that
lets us explore the scalability of a planner's search and preprocessing
capabilities, and illustrates what we perceive to be one of the main limitations
of current off-the-shelf planners for our applications: they often spend a long
time computing ground instances, even when most of these instances are not
required during plan search.

@@

We now turn our attention to a set of experiments arising from the GIVE domain.
Besides using many of the planners from the previous set of experiments (FF,
Metric-FF, and SGPLAN), we also expand our testing to include the FF(h$_a$)
\citep{keyder08:_ff_h_plann_for_plann}, LAMA \citep{richter08:_lama_plann}, and
C$^3$ \citep{lipovetzky08:_c3} planners. Each of these additional planners
competed in the deterministic ``sequential, satisfying'' track of the 2008
International Planning Competition; all planners performed well on the
competition domains, with LAMA the overall winner of the track.\footnote{See
\texttt{http://ipc.informatik.uni-freiburg.de/} for details of the 2008 IPC.}

@@

Preprocessing (parsing, grounding, etc.) time generally plays less of a role in
GIVE, compared with the sentence generation domain, however, its effects still
contribute significantly to the overall running time of a number of planners.
Figure~\ref{fig:give-minimal-grounding} shows the grounding time for FF, LAMA,
and SGPLAN on the minimal GIVE problems, compared with the total running time.
In the unordered variant of the minimal GIVE domain
(Figure~\ref{fig:give-minimal-grounding}(a)), the grounding time in LAMA and
SGPLAN accounts for a significant fraction of the total runtime: SGPLAN spends
around 40\% of its total runtime on preprocessing; for LAMA, this number
rises to at least 80\% for our test problems. For FF, the preprocessing time is
much less important than the search time, especially for large problem
instances. In the ordered case (Figure~\ref{fig:give-minimal-grounding}(b)), the
actual time spent on preprocessing is essentially unchanged from the unordered
case, and search time dominates the total runtime for all three planners.
Overall, however, FF is now much better at controlling the search, compared with
the other planners and its performance on the unordered variant of the problem.

@@

Results for the $h=20$, $n=10$ case with $w$ ranging from $1$ to $40$ are shown
in Figure~\ref{fig:give-extracells}. As in Experiment~3, a number of planners
again fail to solve all the problems: Metric-FF, FF(h$_a$), and C$^3$ solve only
a few instances, while FF only scales to $w=23$. In the unordered version of the
domain, SGPLAN easily solves inputs beyond $w=40$ in well less than a second.
LAMA is also reasonably successful on these problems, however, its runtimes grow
more quickly than SGPLAN, with LAMA taking almost 5 seconds to solve the $w=40$
problem instance. In the ordered case, we again see behaviour similar to that of
Experiment~3: for the problem instances FF is able to solve, it performs
significantly better than LAMA and SGPLAN. (SGPLAN's long term runtime
appears to be growing at a slower rate than FF's, and so even if FF could be
scaled to larger problem instances, it seems possible that SGPLAN might overtake
FF as the better performer.) However, the overall planning times for most of
these instances are concerning since times over a couple seconds will negatively
affect the response time of an NLG system, which must react in real time to user
actions.

>>>>

We now return to the original question of the paper: is planning technology
ready for realistic applications in natural language generation? To investigate
this question we consider two sets of experiments, designed to evaluate the
performance of several planners on the NLG planning domains from the previous
section. Starting with the CRISP domain, we first present a scenario which
focuses on the generation of referring expressions with a tiny grammar
(Section~\ref{sec:exper-1:-sent}). We then look at a setting in which CRISP is
used for surface realization with the XTAG Grammar \citep{xtag01:_tr}, a
large-scale TAG grammar for English (Section~\ref{sec:exper-2:-sent-xtag}). In
the second set of experiments we investigate the GIVE domain. We begin with a
domain that is similar to the classic Gridworld
(Section~\ref{sec:exper-3:-minim}), and then add extra grid cells to the world
that are not necessary to complete the task
(Section~\ref{sec:experiment-4:-give}). We also investigate the role that goal
ordering plays in these problems. These experiments are configured in a way that
lets us explore the scalability of a planner's search and preprocessing
capabilities, and illustrate what we perceive to be one of the main limitations
of current off-the-shelf planners for our applications: they often spend a long
time computing ground instances, even when most of these instances are not
required during plan search.

@@

We now turn our attention to a set of experiments arising from the GIVE domain.
Besides using many of the planners from the previous set of experiments (FF,
Metric-FF, and SGPLAN), we also expand our testing to include the FF(h$_a$)
\citep{keyder08:_ff_h_plann_for_plann}, LAMA \citep{richter08:_lama_plann}, and
C$^3$ \citep{lipovetzky08:_c3} planners. Each of these additional planners
competed in the deterministic ``sequential, satisficing'' track of the 2008
International Planning Competition; all planners performed well on the
competition domains, with LAMA the overall winner of the track.\footnote{See
\texttt{http://ipc.informatik.uni-freiburg.de/} for details of the 2008 IPC.}

@@

Preprocessing time (parsing, grounding, etc.) generally plays less of a role in
GIVE, compared with the sentence generation domain; however, its effects still
contribute significantly to the overall running time of a number of planners.
Figure~\ref{fig:give-minimal-grounding} shows the grounding time for FF, LAMA,
and SGPLAN on the minimal GIVE problems, compared with the total running time.
In the unordered variant of the minimal GIVE domain
(Figure~\ref{fig:give-minimal-grounding}(a)), the grounding time in LAMA and
SGPLAN accounts for a significant fraction of the total runtime: SGPLAN spends
around 40\% of its total runtime on preprocessing; for LAMA, this number
rises to at least 80\% for our test problems. For FF, the preprocessing time is
much less important than the search time, especially for large problem
instances. In the ordered case (Figure~\ref{fig:give-minimal-grounding}(b)), the
actual time spent on preprocessing is essentially unchanged from the unordered
case, and search time dominates the total runtime for all three planners.
Overall, however, FF is now much better at controlling the search, compared with
the other planners and its performance on the unordered variant of the problem.

@@

Results for the $h=20$, $n=10$ case with $w$ ranging from $1$ to $40$ are shown
in Figure~\ref{fig:give-extracells}. As in Experiment~3, a number of planners
again fail to solve all the problems: Metric-FF, FF(h$_a$), and C$^3$ solve only
a few instances, while FF only scales to $w=23$. In the unordered version of the
domain, SGPLAN easily solves inputs beyond $w=40$ in well less than a second.
LAMA is also reasonably successful on these problems; however, its runtimes grow
more quickly than SGPLAN, with LAMA taking almost 5 seconds to solve the $w=40$
problem instance. In the ordered case, we again see behaviour similar to that of
Experiment~3: for the problem instances FF is able to solve, it performs
significantly better than LAMA and SGPLAN. (SGPLAN's long term runtime
appears to be growing at a slower rate than FF's, and so even if FF could be
scaled to larger problem instances, it seems possible that SGPLAN might overtake
FF as the better performer.) However, the overall planning times for most of
these instances are concerning since times over a couple seconds will negatively
affect the response time of an NLG system, which must react in real time to user
actions.

### 2009-07-30 tags/ant-crisp/Papers/ci-spark/nlg-as-planning.tex
Tagged the last version of CRISP with Ant buildfile and directory structure before converting to Maven.

Viewing generation as a planning problem has a long tradition in the NLG
literature. \citet{perrault80} presented an approach to discourse planning in
which the planning operators represented individual speech acts such as
``request'' and ``inform''. This idea was later expanded, e.g., by
\citet{young94dpocl}. On the other hand, researchers such as
\citet{appelt:planning} and \citet{hovy88} used techniques from hierarchical
planning to expand a high-level plan consisting of speech acts into more
detailed specifications of individual sentences. Their systems also covered some
aspects of sentence planning. However, these systems used very expressive
logics, which were designed to reason about beliefs and intentions, to represent
the planning state and the planning operators. Furthermore, most of them used
ad-hoc planning algorithms with rather naive search strategies, which did not
scale to realistic inputs. As a consequence, the NLG-as-planning approach was
mostly marginalized throughout the 1990s.

More recently, there has been a string of publications by various authors with a
renewed interest in the generation-as-planning approach, motivated by the
ongoing development of increasingly more efficient and expressive planners. For
instance, \cite{KolSto07} propose an approach to sentence generation (i.e., the
sentence planning and surface realization modules of the pipeline) as
planning---an approach we will explore in more detail below
(Section~\ref{sec:domain-crisp}). \citet{Steedman-Petrick:07} revisit the
analysis of indirect speech acts with modern planning technology, viewing the
problem as an instance of the general problem of planning with incomplete
information and sensing actions. In addition, \citet{benotti08b} uses planning
to explain the accommodation of presuppositions, and
\citet{brenner08:_contin_multiag_plann_approac_to_situat_dialog} use multi-agent
planning to model the joint problem-solving behaviour of agents in a situated
dialogue. While these approaches focus on different issues compared to the
1980's NLG-as-planning literature, they all apply \emph{existing},
well-understood planning approaches to linguistic problems, in order to utilise
the rich set of modelling tools provided by modern planners, and in the hope
that such planners can efficiently solve the hard search problems that arise in
NLG \citep{KolStr02}. This paper aims to investigate whether existing planners
achieve this latter goal.

>>>>

Viewing generation as a planning problem has a long tradition in the NLG
literature. \citet{perrault80} presented an approach to discourse planning
in which the planning operators represented individual speech acts such as
``request'' and ``inform''. This idea was later expanded, e.g., by
\citet{young94dpocl}. On the other hand, researchers such as
\citet{appelt:planning} and \citet{hovy88} used techniques from
hierarchical planning to expand a high-level plan consisting of speech acts
into more detailed specifications of individual sentences. Although these
systems covered some aspects of sentence planning, they also used very
expressive logics designed to reason about beliefs and intentions, in order
to represent the planning state and the planning operators. Most of these
systems also used ad-hoc planning algorithms with rather naive search
strategies, which did not scale well to realistic inputs. As a consequence,
the NLG-as-planning approach was mostly marginalized throughout the 1990s.

More recently, there has been a string of publications by various authors
with a renewed interest in the generation-as-planning approach, motivated
by the ongoing development of increasingly more efficient and expressive
planners. For instance, \cite{KolSto07} propose an approach to sentence
generation (i.e., the sentence planning and surface realization modules of
the pipeline) as planning---an approach we explore in more detail below
(Section~\ref{sec:domain-crisp}). \citet{Steedman-Petrick:07} revisit the
analysis of indirect speech acts with modern planning technology, viewing
the problem as an instance of planning with incomplete information and
sensing actions. In addition, \citet{benotti08b} uses planning to explain
the accommodation of presuppositions, and
\citet{brenner08:_contin_multiag_plann_approac_to_situat_dialog} use
multi-agent planning to model the joint problem solving behaviour of agents
in a situated dialogue. While these approaches focus on different issues
compared to the 1980's NLG-as-planning literature, they all apply
\emph{existing}, well-understood planning approaches to linguistic
problems, in order to utilise the rich set of modelling tools provided by
modern planners, and in the hope that such planners can efficiently solve
the hard search problems that arise in NLG \citep{KolStr02}. This paper
aims to investigate whether existing planners achieve this latter goal.

### 2009-07-30 tags/ant-crisp/Papers/dl-gre-jair/conclusion.tex
Tagged the last version of CRISP with Ant buildfile and directory structure before converting to Maven.

In this paper, we have explored the idea of viewing the generation of
singular REs as the problem of computing a DL formula with a given
extension.  We have shown how such formulas can be computed
efficiently (for \alc\ and \el) by adapting existing algorithms from
the literature.  The \el\ algorithm is able to generate 95\% of the
non-relational and 67\% of the relational REs from
\newcite{viethen06:_algor_for_gener_refer_expres}.  Both algorithms
are extremely efficient (350 ms and 140 ms respectively to generate
relational REs for all individuals in a random model with 100
individuals); to our knowledge, these are by far the fastest runtimes
for relational GRE reported in the literature.  We have made our
implementation available online at
\url{http://code.google.com/p/crisp-nlg/wiki/DlGre}.

Because they compute referring expressions for all individuals in the
domain at once, our algorithms will perform especially strongly in
static settings, such as the generation of descriptions for museum
exhibits, in which the individuals and their properties don't change
much.  However, even in more dynamic settings, our algorithms have a
chance to outperform search algorithms like Dale and Haddock's in the
average case because they can't get stuck in unproductive branches of
the search space. Nevertheless, one interesting question for future
research is how to incrementally update simulation classes when the
model changes. Similarly, it would be interesting to explore how
linguistic constraints and different attribute orderings can be taken
into account without slowing the algorithms down, how our algorithms
could be integrated with more standard DL T-Box inferences, and how the
algorithms can be adapted to use inverse relations or to compute REs
for sets. In exploring these extensions we will be able to draw on a
rich body of literature that considers many variants of simulation
algorithms.

@@

\begin{small}
\paragraph{Acknowledgments.} We are grateful to Hector Geffner (who
independently suggested to view GRE as computation of DL formulas),
Kees van Deemter, and Emiel Krahmer for interesting discussions.  We
also thank Jette Viethen and Robert Dale for making their corpus
available, and the reviewers for their comments.
\end{small}

>>>>

In this paper, we have explored the idea of viewing the generation of
singular REs as the problem of computing a DL formula with a given
extension.  We have shown how such formulas can be computed
efficiently (for \alc\ and \el) by adapting existing algorithms from
the literature.  The \el\ algorithm is able to generate 95\% of the
non-relational and 67\% of the relational REs from
\citeA{viethen06:_algor_for_gener_refer_expres}.  Both algorithms
are extremely efficient (350 ms and 140 ms respectively to generate
relational REs for all individuals in a random model with 100
individuals); to our knowledge, these are by far the fastest runtimes
for relational GRE reported in the literature.  We have made our
implementation available online at
\url{http://code.google.com/p/crisp-nlg/wiki/DlGre}.

Because they compute referring expressions for all individuals in the
domain at once, our algorithms will perform especially strongly in
static settings, such as the generation of descriptions for museum
exhibits, in which the individuals and their properties don't change
much.  However, even in more dynamic settings, our algorithms have a
chance to outperform search algorithms like Dale and Haddock's in the
average case because they can't get stuck in unproductive branches of
the search space. Nevertheless, one interesting question for future
research is how to incrementally update simulation classes when the
model changes. Similarly, it would be interesting to explore how
different linguistic constraints and attribute orderings can be taken
into account efficiently, how our algorithms
could be integrated with more standard DL T-Box inferences, and how they
 can be adapted to use inverse relations or to compute REs
for sets. In exploring these extensions we will be able to draw on a
rich body of literature that has already considered many variants of simulation
algorithms addressing similar issues.

@@

### 2009-07-30 tags/ant-crisp/Papers/dl-gre-jair/introduction.tex
Tagged the last version of CRISP with Ant buildfile and directory structure before converting to Maven.

Referring expressions (REs) involving relations, in particular, have
received increasing attention recently, specially in the context of
spatial referring expressions in situated generation (e.g.\
\cite{kelleher06:_increm_gener_of_spatial_refer}), where it seems
particularly natural to use expressions such as ``the book on the
table''.  However, the classical algorithm by
\newcite{dale91:_gener_refer_expres_invol_relat} was recently shown to
be unable to generate satisfying REs in practice
\cite{viethen06:_algor_for_gener_refer_expres}.  Furthermore, the Dale
and Haddock algorithm and most of its successors (such as
\cite{kelleher06:_increm_gener_of_spatial_refer}) are vulnerable to
the problem of ``infinite regress'', where the algorithm jumps back
and forth between generating descriptions for two related individuals
infinitely, as in ``the book on the table which supports a book on the
table \ldots''.

In this paper, we propose to view GRE as the problem of computing a
formula of description logic (DL) that denotes exactly the set of
individuals that we want to refer to.  This very natural idea has been
mentioned in passing before
\cite{Krahmer2003,gardent07:_gener_bridg_defin_descr}; however, we
take it one step further by proposing DL as an interlingua for
comparing the REs produced by different approaches to GRE.  In this
way, we can organize existing GRE approaches in an expressiveness
hierarchy.  For instance, the classical Dale and Reiter algorithms
compute purely conjunctive formulas;
\newcite{deemter02:_gener_refer_expres} extends this language by
adding the other propositional connectives, whereas
\newcite{dale91:_gener_refer_expres_invol_relat} extends it by
allowing existential quantification.

Furthermore, the view of GRE as a problem of computing DL formulas
with a given extension allows us to apply existing algorithms for the
latter problem to obtain efficient algorithms for GRE.  We present
algorithms that compute such formulas for the description logics \el\
(which allows only conjunction and existential quantification) and
\alc\ (which also allows negation).  These algorithms effectively
compute REs for all individuals in the domain at the same time, which
allows them to systematically avoid the infinite regress problem.  The
\el\ algorithm is capable of generating 67\% of the relational REs in
the \newcite{viethen06:_algor_for_gener_refer_expres} dataset, in about
15 milliseconds.  The \alc\ algorithm is even faster; it computes
relational REs for all 100 individuals in a random model in 140
milliseconds.

>>>>

Referring expressions (REs) involving relations, in particular, have
received increasing attention recently; especially in the context of
spatial referring expressions in situated generation
\cite<e.g.,>{kelleher06:_increm_gener_of_spatial_refer}, where it seems
particularly natural to use expressions such as ``the book on the
table''.  However, the classical algorithm by
\citeA{dale91:_gener_refer_expres_invol_relat} was recently shown to
be unable to generate satisfying REs in practice
\cite{viethen06:_algor_for_gener_refer_expres}.  Furthermore, the Dale
and Haddock algorithm and most of its successors \cite<such
as>{kelleher06:_increm_gener_of_spatial_refer} are vulnerable to the
problem of ``infinite regress'', where the algorithm jumps back and
forth between generating descriptions for two related individuals
infinitely, as in ``the book on the table which supports a book on the
table \ldots''.

In this paper, we propose to view GRE as the problem of computing a
formula of description logic (DL) that denotes exactly the set of
individuals that we want to refer to.  This very natural idea has been
mentioned in passing before
\cite{Krahmer2003,gardent07:_gener_bridg_defin_descr}; however, we
take it one step further by proposing DL as an interlingua for
comparing the REs produced by different approaches to GRE.  In this
way, we can organize existing GRE approaches in an expressiveness
hierarchy.  For instance, the classical Dale and Reiter algorithms
compute purely conjunctive formulas;
\citeA{deemter02:_gener_refer_expres} extends this language by
adding the other propositional connectives, whereas
\citeA{dale91:_gener_refer_expres_invol_relat} extends it by
allowing existential quantification.

Furthermore, the view of GRE as a problem of computing DL formulas
with a given extension allows us to apply existing algorithms for the
latter problem to obtain efficient algorithms for GRE.  We present
algorithms that compute such formulas for the description logics \el\
(which allows only conjunction and existential quantification) and
\alc\ (which also allows negation).  These algorithms effectively
compute REs for all individuals in the domain at the same time, which
allows them to systematically avoid the infinite regress problem.  The
\el\ algorithm is capable of generating 67\% of the relational REs in
the \citeA{viethen06:_algor_for_gener_refer_expres} dataset, in about
15 milliseconds.  The \alc\ algorithm is even faster; it computes
relational REs for all 100 individuals in a random model in 140
milliseconds.

### 2009-07-30 tags/ant-crisp/Papers/experiences/conclusion.tex
Tagged the last version of CRISP with Ant buildfile and directory structure before converting to Maven.

The NLG community's recent interest in planning presents a valuable
opportunity for planning researchers.  While GIVE highlights
planning-related challenges such as plan execution monitoring and plan
presentation (i.e., summarisation and elaboration), it also offers a
platform on which such technologies can be evaluated in experiments with
human users.  Furthermore, more mainstream research on reasoning under
uncertainty, planning with incomplete information and sensing, and resource
management can also be applied in this domain setting. Provided some of the
challenges we have highlighted can be addressed, projects like GIVE (and
other NLG-inspired problem domains) offer a constructive platform for the
planning community to showcase their techniques to a wider audience---and
to improve the quality of their tools for real-world planning tasks.

>>>>

The NLG community's recent interest in planning presents a valuable
opportunity for planning researchers.  While GIVE highlights
certain challenges such as plan execution monitoring and plan presentation
(i.e., summarisation and elaboration), it also offers a platform on which
such technologies can be evaluated in experiments with human users.
Furthermore, research related to reasoning under uncertainty, resource
management, and planning with knowledge and sensing, can also be
investigated in this domain setting. Provided some of the challenges we
have highlighted can be addressed, projects like GIVE (and other
NLG-inspired problem domains) offer a constructive platform for the
planning community to showcase their techniques to a wider audience---and
to improve the quality of their tools for real-world planning tasks.

### 2009-07-30 tags/ant-crisp/Papers/experiences/domains.tex
Tagged the last version of CRISP with Ant buildfile and directory structure before converting to Maven.

The GIVE task comes with a number of interesting research challenges
both for the NLG and the planning communities.  On the NLG side, one
major problem is that the natural language instruction sequences can't
simply be obtained by verbalising each action instance in a plan in turn.
For instance, the NLG system could generate an instruction sequence
starting with ``turn left; walk forward; press the button; turn right; walk
forward; walk forward; turn right; walk forward; walk forward; turn left;
walk forward''.  However, such directions would be clumsy and boring, and
it is to be expected that frustrated users will quickly cancel an
interaction with such a system, resulting in a bad evaluation score.
Having just pressed the button, it would be much better to say ``turn
around and walk through the door''. That is, plans must be
\emph{summarised} by merging multiple action instances into a single
instruction, in order to present them to the
user.\footnote{This task is similar to the construction of ``macro''
actions, but is primarily intended for plan presentation.}
Conversely, it may also be necessary to \emph{elaborate} on a single
planning step by expressing it with several instructions.  Having just
entered the top left room, it may be easier for the user to understand the
instruction ``walk to the centre of the room; turn right; now press the
green button in front of you'', rather than the instruction ``press the
green button on the wall to your right''.  To plan the referring expression
``the green button in front of you'' at a time when the button is not yet
in front of the user, the NLG system must keep track of the hypothetical
changes in what objects are visible to the user.  Thus the acts of
referring and instructing, and the modules for planning and NLG, must be
tightly integrated.

From a planning perspective, GIVE imposes very strict runtime requirements
on the planner and its associated modules: planning must happen in real
time and the system must respond to a user in a timely fashion.  If the
system takes too much time deliberating over an instruction to give, rather
than actually giving this instruction, the user may have walked or turned
away, thus making the instruction invalid.  Furthermore, plan execution
monitoring also plays an important role in the GIVE problem.  At a high
level, the system needs to monitor a user's actions and compare them
against the generated instruction set to determine if the user has
correctly followed directions or not. In the case of the latter, new
instructions may have to be generated. In practice, the situation can be
quite complicated since the mental state of the user is not known and so
the system must observe the user's actions in real time. For example, a
user directed to ``turn around and walk through the door'' may not
necessarily perform these actions to the letter, i.e., immediately turning
180 degrees and proceeding directly to the door. Instead, the user might
take a roundabout route through the room, eventually exiting out the door.
Although the user's actions do not match the generated instructions
exactly, they meet the intended goal. The system must be able to identify
such ``equivalent'' plans and not immediately generate new instructions as
soon as the user's actions have gone off course. Furthermore, a user can
communicate certain intentions to the system, both through action and
inaction. For instance, the system should infer that a user has failed to
follow instructions if the user exits a room when given a directive to
``walk to the centre of the room''. The system should also make a
similar conclusion if a user simply does nothing when given the
instruction.

>>>>

The GIVE task comes with a number of interesting research challenges
both for the NLG and the planning communities.  On the NLG side, one
major problem is that verbalising each action instance in the plan in
turn makes for very poor natural-language instructions.  For instance,
the NLG system could generate an instruction sequence starting with
``turn left; walk forward; press the button; turn right; walk forward;
walk forward; turn right; walk forward; walk forward; turn left; walk
forward''.  However, such directions would be clumsy and boring, and
it is to be expected that frustrated users will quickly cancel an
interaction with such a system, resulting in a bad evaluation score.
Having just pressed the button, it would be much better to say ``turn
around and walk through the door''. That is, plans must be
\emph{summarised} by merging multiple action instances into a single
instruction, in order to present them to the user.\footnote{This task
  is similar to the construction of ``macro'' actions, but is
  primarily intended for plan presentation.}  Conversely, it may also
be necessary to \emph{elaborate} on a single planning step by
expressing it with several instructions.  Having just entered the top
left room, it may be easier for the user to understand the instruction
``walk to the centre of the room; turn right; now press the green
button in front of you'', rather than the instruction ``press the
green button on the wall to your right''.  To plan the referring
expression ``the green button in front of you'' at a time when the
button is not yet in front of the user, the NLG system must keep track
of the hypothetical changes in what objects are visible to the user.
Thus the acts of referring and instructing, and the modules for
planning and NLG, must be tightly integrated.

From a planning perspective, GIVE imposes very strict runtime
requirements on the planner and its associated modules: planning must
happen in real time and the system must respond to a user in a timely
fashion.  If the system takes too much time deliberating over an
instruction to give, rather than actually giving this instruction, the
user may have walked or turned away, thus making the instruction
invalid.  Furthermore, plan execution monitoring also plays an
important role in the GIVE problem.  At a high level, the system needs
to monitor a user's actions and compare them against the generated
instruction set to determine if the user has correctly followed
directions or not. In the case of the latter, new instructions may
have to be generated. In practice, the situation can be quite
complicated since the mental state of the user is not known and so the
system must be inferred from observing the user's actions in real
time. For example, a user directed to ``turn around and walk through
the door'' may not necessarily perform these actions to the letter,
i.e., immediately turning 180 degrees and proceeding directly to the
door. Instead, the user might take a roundabout route through the
room, eventually exiting out the door.  Although the user's actions do
not match the generated instructions exactly, they meet the intended
goal. The system must be able to identify such ``equivalent'' plans
and not immediately generate new instructions as soon as the user's
actions have gone off course. Furthermore, a user can communicate
certain intentions to the system, both through action and
inaction. For instance, the system should infer that a user has failed
to follow instructions if the user exits a room when given a directive
to ``walk to the centre of the room''. The system should also make a
similar conclusion if a user simply does nothing when given the
instruction.

### 2009-07-30 tags/ant-crisp/Papers/rmrs-08/rmrs.tex
Tagged the last version of CRISP with Ant buildfile and directory structure before converting to Maven.

The {\em satisfaction} relation between an \rmrs\ $\varphi$ and a
finite constructor tree $\tau$ is defined in terms of several
assignment functions.  First, a node variable assignment
function $\alpha: \nvar \rightarrow D(\tau)$ maps the node
variables in an \rmrs\ to the nodes of $\tau$.  Second, a
base language assignment function $g:\ovar \rightarrow \Sigma_0$ 
maps the base variables to nullary constructors
representing variables in the base language. Finally, a
function $\sigma$ from $\pred$ to the power set of $\Sigma_{\geq 1}$ maps
each \rmrs\ predicate symbol to a set of constructors from $\Sigma$.
As we'll see shortly, this relation allows an \rmrs\ to underspecify lexical
ambiguities.

>>>>

The {\em satisfaction} relation between an \rmrs\ $\varphi$ and a
finite constructor tree $\tau$ is defined in terms of several
assignment functions.  First, a node variable assignment function
$\alpha: \nvar \rightarrow D(\tau)$ maps the node variables in an
\rmrs\ to the nodes of $\tau$.  Second, a base language assignment
function $g:\ovar \rightarrow \Sigma_0$ maps the base variables to
nullary constructors representing variables in the base
language. Finally, a function $\sigma$ from $\pred$ to the power set
of $\Sigma_{\geq 1}$ maps each \rmrs\ predicate symbol to a set of
constructors from $\Sigma$.  As we'll see shortly, this function
allows an \rmrs\ to underspecify lexical ambiguities.
